Do return series have power-law tails with the same index?

### 2017-04-14 papers/FX/draft.tex
Revised the script.
A major revision of the FX manuscript.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
\begin{eqnarray*}
  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
    0, \left( {K \over \alpha}  \right)^2
  \right) \\
  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
    0, \left( {K^{2\alpha}}  \right)^2
  \right)
\end{eqnarray*}
where the tail index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of the estimate of $K$
and $K^\alpha$ can be constructed. 
Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd column of
figure~\ref{fig:sectors_parameters}.

Secondly, one can see that the values of $\alpha$ and $K$ are appear
positively correlated. This prompts Pareto tails of the return
distributions. We expain more about this in
section~\ref{sec:pareto_tail}.

\subsection{A test of equal tail indices based on Hill estimator}
From \eqref{eq:2} one can deduce
\begin{equation}
  \label{eq:x1}
  \sqrt k (\hat \alpha_A - \hat \alpha_B) \overset{d}{\to} N(0, \alpha_A^2 - \alpha_B^2)  
\end{equation}
where $\hat \alpha_A$ and $\hat \alpha_B$ are Hill estimates of
$\alpha_A$ and $\alpha_B$, the tail indices of two series $A$ and
$B$. The weak convergence of \eqref{eq:x1} allows one to construct a
test of the equality of $\alpha_A$ and $\alpha_B$.

@@

``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:Hoga_pair}.

@@

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Energy_Single.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Consumer_Staples_Single.pdf}
  \end{minipage}
  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
by simulations from the limit \ds .}
  \label{fig:Hoga_Single}
\end{figure}

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Energy_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 1cm, 1cm, 1cm}, clip
    ]{Hoga_Consumer_Staples_pair.pdf}
  \end{minipage}
  \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
    ``Consumer Staples'' sectors of
    S\&P 500. The green, blue and red colors represent, respectively, the cases when  
    the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
    Grey points stand for pairs for which $T_n$ is below
    the 85\% \asy\ quantile;
    black points represent pairs for which
    the computation of $T_n$ fails for given precision
    requirements and time limits.}
  \label{fig:Hoga_pair}
\end{figure}

\par
To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size $n=1304$ as the S\&P 500 series. 
Both pieces are $t$-distributed with distinct degrees
of freedom. The results are shown in Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the larger the minimum
tail index in the concatenated pair.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth, trim={1.0cm 1.5cm 1cm 2cm},
  clip]{t_sim_pair.pdf}
  \caption{Test  of concatenated $t$-samples with different
    degrees of freedom. Numbers on the axes are the degrees of freedom in the subsamples. For an interpretation of the colored
bullets, see Figure~\ref{fig:Hoga_pair}. The graph shows the limited power of the test. In particular, if
both degrees of freedom are relatively large it loses the capablity of  distinguishing the \ds .}
  \label{fig:t_sim_pair}
\end{figure}
\par
A major problem in this test is the \asy\ distribution of the test statistic under the
null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
To find out about this problem we compared the \ds s of $T_n$ 
with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom and the limit \ds\ of $T_n$.
The estimated density functions are shown in Figure~\ref{fig:Q1_distr}.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \caption{{\red Change $Q_1$ on $x$-axis} Comparison of the \asy\ distribution of the test statistic $T_n$ in \eqref{eq:4} under the null hypothesis 
and the \ds\ of $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees of freedom.}
  \label{fig:Q1_distr}
\end{figure}
As seen in the graph, the asymptotic distribution 
assigns significantly more mass to the tail than the \ds s of $T_n$ do. For comparison, we list a few quantiles of these \ds s in
Table~\ref{tab:HogaAsymptotic}, showing major differences between the \asy\ and finite-sample \ds s.
\begin{table}[htb!]
  \centering
  \begin{tabular}{l|c|c|c|r}
    & \multicolumn{4}{c}{Quantiles} \\[2mm]
    \hline
    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
    \hline
    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
  \end{tabular}
  \caption{Quantiles of the test statistic $T_n$ for $n=1304$ $t$-distributed samples with 3 and 4 degrees of freedom as well
as the corresponding quantiles for the limiting \ds\ of $T_n$. In particular, there are huge differences between the three \ds s  
for the higher quantiles. 
    }
  \label{tab:HogaAsymptotic}
\end{table}

@@

\subsection{An utility based argument for equal tail indices} \setcounter{equation}{0}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
  meaning of these constants?}
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

@@

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
    {\red Is this worth a theorem?}
    {\blue No}
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
\end{remark}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
    (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other
      figures, and use the same structure of the formulas.}
  }
  \label{fig:phi_hat_pareto}
\end{figure}

@@

\begin{lemma}
  \label{lemma:II}
  There exists $x_0 > 0$ such that ${\pd f \over \pd \nu}(x_0, \nu) = 0$ and
  $\forall x \in (0, x_0), {\pd f \over \pd \nu}(x, \nu) > 0$ and
  $\forall x \in (x_0, \infty), {\pd f \over \pd \nu}(x, \nu) < 0$.
\end{lemma}

>>>>

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
\begin{eqnarray*}
  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
    0, \left( {K \over \alpha}  \right)^2
  \right) \\
  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
    0, \left( {K^{2\alpha}}  \right)^2
  \right)
\end{eqnarray*}
where the tail index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of the estimate of $K$
and $K^\alpha$ can be constructed. 
Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
    of the S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals. The confidence bands of the
    corresponding Hill estimates $\hat \alpha$ of these sectors are
    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
    instead of $K^\alpha$.
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd column of
figure~\ref{fig:sectors_parameters}.

Secondly, one can see that the values of $\alpha$ and $K$ appear
positively correlated. This prompts Pareto tails of the return
distributions. We expain more about this in
section~\ref{sec:pareto_tail}.

Thirdly, figure~\ref{fig:sectors_parameters} shows the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are on
average larger than those in the ``Consumer Staples'' sector. A larger
value of $K$ means, for a given value of $p$, i.e. a given probability
of losses, large losses are more probable. Thus one can conclude that
these two sectors are considerablly riskier than the ``Consumer
Staples'' sector. This is of course a confirmation of one's economic
instinct.

Yet another indication from figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the correlation between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in the last section, moving along
a curve of equal preference in the direction of increasing $\alpha$, $K$ also
increases. So the strong positive correlation seen in the ``Energy''
sector suggests these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

\subsection{A test of equal tail indices based on Hill estimator}
Suppose we have two weakly dependent series $X_1, \dots, X_n \sim F_X$ and
$Y_1, \dots, Y_n \sim F_Y$. Both $F_X$ and $F_Y$ have power-law tails
with indices $\alpha_X$ and $\alpha_Y$, respectively. Assume also
$X_i$ is independent of $Y_j$ for all $1 \leq i \leq j \leq n$.
From \eqref{eq:2} one can deduce
\begin{equation}
  \label{eq:x3}
  \sqrt k
  \begin{pmatrix}
    {\hat \alpha_X - \alpha_X} \\
    {\hat \alpha_X - \alpha_X} \\
  \end{pmatrix} \overset{d}{\to}
  \begin{pmatrix}
    Z_X \\
    Z_Y
  \end{pmatrix}
  \sim
  N\left(
    0, \text{diag}(\alpha_X^2, \alpha_Y^2)
\end{equation}
where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimates of
$\alpha_X$ and $\alpha_Y$. Then it follows from continuous mapping
theorem
\begin{equation}
  \label{eq:x1}
  \sqrt k [(\hat \alpha_X -\alpha_X) - (\hat \alpha_Y - \alpha_Y)]
  \overset{d}{\to}
  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)
\end{equation}
The weak convergence of \eqref{eq:x1} allows one to construct a test
of the equality $\alpha_X = \alpha_Y$. As the null hypothesis, assume
$\alpha_X = \alpha_Y$. Under this hypothesis, \eqref{eq:x1} means the
test statistic $\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0,
\alpha_X^2 + \alpha_Y^2)$. 
We apply this test to the equities in the ``Energy'',
``Consumder Staples'' and ``Information Technology''
sectors of the S\&P 500 index. The results are shown in
\ref{fig:PairTest}.
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_Energy.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_CS.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_IT.pdf}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_Energy_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_CS_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_IT_pair.pdf}
  \caption{
    {\em 1st row}: Test for changing tail index using Hill estimator based on
    concatenated series in the ``Energy'' and ``Consumer Staples''
    sectors of S\&P 500. The green, blue and red colors represent,
    respectively, the cases when the 85\%, 90\% and 95\% quantiles of
    the limit distribution of the test statistic is exceeded. 
    Grey points stand for pairs for which the test statistic is below
    the 85\% \asy\ quantile;
    {\em 2nd row:} Test for changing tail index or scale parameter
    using Hoga's test based on the same concatenated series. black points represent
    pairs for which the computation of $T_n$ fails for given precision
    requirements and time limits.
    The same number of upper-order statistics are used for both tests.
  }
  \label{fig:PairTest} 
One can see from the figure that tail indices of equities in the
``Energy'' or ``Information Technology'' sectors vary more than do
their counterparts in the ``Consumer Staples'' sector, as the null
hypothesis is rejected more often for members of these two sectors.
Moreover, these figures suggest that the test based on Hill estimator
is more powerful in distinguishing differences between tail indices,
as it results in more rejections for the ``Energy'' and the
``Information Technology'' sectors than does the other test. We
argue this can be the result of slow convergence of the distribution
of the test statistic to its asymptotic distribution. We elaborate on
this in the section~\ref{sec:Hoga}.

Nevertheless, one should to bear in mind that the weak convergence
\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
series are independent of each other, which is generally untrue for
two return series in the same market.

@@

``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:PairTest}.

@@

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Energy_Single.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_CS_Single.pdf}
  \end{minipage}
  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
by simulations from the limit \ds .}
  \label{fig:Hoga_Single}
\end{figure}

% \begin{figure}[htb!]
%   \begin{minipage}{0.5\linewidth}
%     \includegraphics[
%       width=\textwidth,
%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
%     ]{Hoga_Energy_pair.pdf}
%   \end{minipage}\hfill
%   \begin{minipage}{0.5\linewidth}
%     \includegraphics[
%       width=\textwidth,
%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
%     ]{Hoga_CS_pair.pdf}
%   \end{minipage}
%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
%     ``Consumer Staples'' sectors of
%     S\&P 500. The green, blue and red colors represent, respectively, the cases when  
%     the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
%     Grey points stand for pairs for which $T_n$ is below
%     the 85\% \asy\ quantile;
%     black points represent pairs for which
%     the computation of $T_n$ fails for given precision
%     requirements and time limits.}
%   \label{fig:Hoga_pair}
% \end{figure}

\par
To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom. Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capablity of distinguishing the \ds\ .
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

\par
A major problem in this test is the \asy\ distribution of the test statistic under the
null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
To find out about this problem we compared the \ds s of $T_n$ 
with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom and the limit \ds\ of $T_n$.
The estimated density functions are shown in on the right of Figure~\ref{fig:t_sim_pair}.
As seen in the graph, the asymptotic distribution assigns
significantly more mass to the tail than the \ds s of $T_n$ do. For
comparison, we list a few quantiles of these \ds s in
Table~\ref{tab:HogaAsymptotic}, showing major differences between the
\asy\ and finite-sample \ds s.
\begin{table}[htb!]
  \centering
  \begin{tabular}{l|c|c|c|r}
    & \multicolumn{4}{c}{Quantiles} \\[2mm]
    \hline
    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
    \hline
    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
  \end{tabular}
  \caption{Quantiles of the test statistic $T_n$ for $n=1304$ $t$-distributed samples with 3 and 4 degrees of freedom as well
as the corresponding quantiles for the limiting \ds\ of $T_n$. In particular, there are huge differences between the three \ds s  
for the higher quantiles. 
    }
  \label{tab:HogaAsymptotic}
\end{table}

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
  meaning of these constants?}
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation in this case
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

@@

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
    {\red Is this worth a theorem?}
    {\blue No, let's change it to a remark.}
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\end{remark}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
    (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other
      figures, and use the same structure of the formulas.}
  }
  \label{fig:phi_hat_pareto}
\end{figure}

@@
